---
title: "Portfolio 4"
---

> In this project, I use the Waves 1 and 2 of the National Longitudinal Study of Adolescent to Adult Health (Add Health) to test the following hypothesis:

> It is hypothesized that young adolescents’ sexual risk-taking (Motivations to Engage in Risky Behaviors, T1) will be associated with higher levels of substance use (Tobacco, Alcohol, Drugs, T1; total effect). Additionally, this effect will persist after controlling for (Parents’ Attitudes, T1; partial effect). However, the relationship between sexual risk-taking and substance use will be stronger among boys than girls (T1; moderated effect).

> First, load the package and data.

```{r load_package&data}
library(haven)

wave1 <- read_spss("wave1.sav")
```

> Next, I'll recode the variables needed for the testing of the hypothesis.

```{r recode}
### Motivation T1
  # recode H1MO1
  wave1$mot1[wave1$H1MO1 == 1] <- 2
  wave1$mot1[wave1$H1MO1 == 2] <- 1
  wave1$mot1[wave1$H1MO1 == 3] <- 0
  wave1$mot1[wave1$H1MO1 == 4] <- -1
  wave1$mot1[wave1$H1MO1 == 5] <- -2
  
  # recode H1MO2
  wave1$mot2[wave1$H1MO2 == 1] <- -2
  wave1$mot2[wave1$H1MO2 == 2] <- -1
  wave1$mot2[wave1$H1MO2 == 3] <- 0
  wave1$mot2[wave1$H1MO2 == 4] <- 1
  wave1$mot2[wave1$H1MO2 == 5] <- 2
  
  # recode H1MO3
  wave1$mot3[wave1$H1MO3 == 1] <- -2
  wave1$mot3[wave1$H1MO3 == 2] <- -1
  wave1$mot3[wave1$H1MO3 == 3] <- 0
  wave1$mot3[wave1$H1MO3 == 4] <- 1
  wave1$mot3[wave1$H1MO3 == 5] <- 2
  
  # recode H1MO5
  wave1$mot5[wave1$H1MO5 == 1] <- 2
  wave1$mot5[wave1$H1MO5 == 2] <- 1
  wave1$mot5[wave1$H1MO5 == 3] <- 0
  wave1$mot5[wave1$H1MO5 == 4] <- -1
  wave1$mot5[wave1$H1MO5 == 5] <- -2

  # recode H1MO6
  wave1$mot6[wave1$H1MO6 == 1] <- 2
  wave1$mot6[wave1$H1MO6 == 2] <- 1
  wave1$mot6[wave1$H1MO6 == 3] <- 0
  wave1$mot6[wave1$H1MO6 == 4] <- -1
  wave1$mot6[wave1$H1MO6 == 5] <- -2
  
  # recode H1MO7
  wave1$mot7[wave1$H1MO7 == 1] <- 2
  wave1$mot7[wave1$H1MO7 == 2] <- 1
  wave1$mot7[wave1$H1MO7 == 3] <- 0
  wave1$mot7[wave1$H1MO7 == 4] <- -1
  wave1$mot7[wave1$H1MO7 == 5] <- -2

  # recode H1MO8
  wave1$mot8[wave1$H1MO8 == 1] <- 2
  wave1$mot8[wave1$H1MO8 == 2] <- 1
  wave1$mot8[wave1$H1MO8 == 3] <- 0
  wave1$mot8[wave1$H1MO8 == 4] <- -1
  wave1$mot8[wave1$H1MO8 == 5] <- -2

  # calculate overall motivation
  wave1$Motivation <- apply(wave1[,c("mot1",
                                         "mot2", 
                                         "mot3", 
                                         "mot5",
                                         "mot6",
                                         "mot7", 
                                         "mot8")], 1, mean)

### Tobacco, Alcohol, Drugs, T1
  wave1$Drug1 <- wave1$H1TO7
  wave1$Drug1[wave1$H1TO7 >95] <- NA
  
### Parents’ Attitudes, T1
  wave1$Mom <- 1
  wave1$Dad <- 1
  
  for(i in 1:nrow(wave1)){
    if(wave1$H1PA1[i] == 7) {
      wave1$Mom[i] <- 0
    }
    
    if(wave1$H1PA4[i] == 7) {
      wave1$Dad[i] <- 0
    }
  }
  
  # recode H1PA1
  wave1$par1[wave1$H1PA1 == 1] <- -2
  wave1$par1[wave1$H1PA1 == 2] <- -1
  wave1$par1[wave1$H1PA1 == 3] <- 0
  wave1$par1[wave1$H1PA1 == 4] <- 1
  wave1$par1[wave1$H1PA1 == 5] <- 2
  
  # recode H1PA2
  wave1$par2[wave1$H1PA2 == 1] <- -2
  wave1$par2[wave1$H1PA2 == 2] <- -1
  wave1$par2[wave1$H1PA2 == 3] <- 0
  wave1$par2[wave1$H1PA2 == 4] <- 1
  wave1$par2[wave1$H1PA2 == 5] <- 2

  # recode H1PA3
  wave1$par3[wave1$H1PA3 == 1] <- -2
  wave1$par3[wave1$H1PA3 == 2] <- -1
  wave1$par3[wave1$H1PA3 == 3] <- 0
  wave1$par3[wave1$H1PA3 == 4] <- 1
  wave1$par3[wave1$H1PA3 == 5] <- 2

  # recode H1PA4
  wave1$par4[wave1$H1PA4 == 1] <- -2
  wave1$par4[wave1$H1PA4 == 2] <- -1
  wave1$par4[wave1$H1PA4 == 3] <- 0
  wave1$par4[wave1$H1PA4 == 4] <- 1
  wave1$par4[wave1$H1PA4 == 5] <- 2
  
  # recode H1PA5
  wave1$par5[wave1$H1PA5 == 1] <- -2
  wave1$par5[wave1$H1PA5 == 2] <- -1
  wave1$par5[wave1$H1PA5 == 3] <- 0
  wave1$par5[wave1$H1PA5 == 4] <- 1
  wave1$par5[wave1$H1PA5 == 5] <- 2
  
  # recode H1PA6
  wave1$par6[wave1$H1PA6 == 1] <- -2
  wave1$par6[wave1$H1PA6 == 2] <- -1
  wave1$par6[wave1$H1PA6 == 3] <- 0
  wave1$par6[wave1$H1PA6 == 4] <- 1
  wave1$par6[wave1$H1PA6 == 5] <- 2
  
  for(i in 1:nrow(wave1)){
    if(wave1$Mom[i] == 0){
      if(wave1$Dad[i] == 0){
        wave1$pAttitude[i] <- NA
      }
      else{
        wave1$pAttitude[i] <- apply(wave1[i,c("par4", 
                                             "par5", 
                                             "par6")], 1, mean)
      }
    }
    else{
      if(wave1$Dad[i] == 0){
        wave1$pAttitude[i] <- apply(wave1[i,c("par1", 
                                             "par2", 
                                             "par3")], 1, mean)
      }
      else{
        wave1$pAttitude[i] <- apply(wave1[i,c("par1", 
                                             "par2", 
                                             "par3", 
                                             "par4", 
                                             "par5", 
                                             "par6")], 1, mean)
      }
    }
  }
  
### Gender, T1 (male = 1; female = 2)
  wave1$BIO_SEX[wave1$BIO_SEX == 6] <- NA
```

> Now I have the appropriate variables, let's build models to test the hypothesis.

```{r stats}
# Hypothesis 1
  H1 <- subset(wave1, complete.cases(wave1[, c("Motivation", 
                                                 "Drug1", 
                                                 "pAttitude", 
                                                 "BIO_SEX")]))
  
    # total effect model
    h1_total_model <- lm(Drug1 ~ Motivation, data = H1)
    summary(h1_total_model)
    
    # partial effect model
    h1_partial_model <- lm(Drug1 ~ Motivation + pAttitude, data = H1)
    summary(h1_partial_model)
    
    # moderated effect model
    h1_moderated_model <- lm(Drug1 ~ Motivation * BIO_SEX + pAttitude, data = H1)
    summary(h1_moderated_model)
```

> The result shows that motivation for risky behaviors was positively associated with substance use. This effect persisted even after accounting for parents’ attitudes. However, the moderation hypothesis was not supported; the data indicated that the relationship between motivation and substance use did not significantly differ between boys and girls. 