---
title: "Portfolio 5"
---

> Adding to portfolio piece 4, this piece tests another hypothesis with the Add Health dataset:

>It is hypothesized that young adolescents who are more engaged in fighting and violence (Fighting and Violence, T1) will engage in more substance use the subsequent year (Tobacco, Alcohol, Drugs, T2; total effect). It is further hypothesized that this association will persist even after controlling for substance use at the previous time point (T1; partial effect).

> First, let's load the package and data.

```{r load_package&data}
library(haven)

wave1 <- read_spss("wave1.sav")
wave2 <- read_spss("wave2.sav")
```

> Next, let's recoed the variables.

```{r recode}
### Tobacco, Alcohol, Drugs, T1
  wave1$Drug1 <- wave1$H1TO7
  wave1$Drug1[wave1$H1TO7 >95] <- NA
  
### Fighting and Violence, T1
  # recode H1FV1
  wave1$H1FV1[wave1$H1FV1 >5] <- NA
  
  # recode H1FV5
  wave1$H1FV5[wave1$H1FV5 >5] <- NA
  
  # recode H1FV7
  wave1$H1FV7[wave1$H1FV7 >5] <- NA
  
  # recode H1FV8
  wave1$H1FV8[wave1$H1FV8 >5] <- NA

  # calculate fighting
  wave1$Fight <- apply(wave1[,c("H1FV1", 
                                "H1FV5", 
                                "H1FV7", 
                                "H1FV8")], 1, sum)
  
### Tobacco, Alcohol, Drugs, T2
  wave2$Drug2 <- wave2$H2TO7
  wave2$Drug2[wave2$H2TO7 > 95] <- NA
```

> Now, let's test the hypothesis.

```{r stats}
### make new data sets
  merged <- merge(wave1, wave2, by = 1, all = TRUE)

 # Hypothesis 2
  H2<- subset(merged, complete.cases(merged[, c("Fight", 
                                                "Drug1", 
                                                "Drug2")]))
  
    # total effect model
    h2_total_model <- lm(Drug2 ~ Fight, data = H2)
    summary(h2_total_model)
    
    # partial effect model
    h2_partial_model <- lm(Drug2 ~ Fight + Drug1, data = H2)
    summary(h2_partial_model)
```

> The result showed a longitudinal relationship between aggression and substance use. Adolescents who engaged in more fighting at Wave 1 were more likely to use substances at Wave 2. This powerful association remained significant even after controlling for their prior levels of substance use at Wave 1, supporting the partial effect hypothesis (β = 0.49, p = .003). This suggests that fighting predicts an increase in substance use over time, independent of initial substance use levels. 