---
title: "Portfolio 6"
---

> This is the last portion of the Add Health analysis. In this piece, I will test a third hypothesis:

> It is hypothesized that neighborhood environment (Neighborhood, T1) will predict lower levels of adolescent depressive symptoms (Feeling Scale, T2; total effect). However, this effect will be attenuated after support from family and friends (Protective Factors, T1) is controlled for (partial effect).

> First, load the package and data.

```{r load_package&data}
library(haven)

wave1 <- read_spss("wave1.sav")
wave2 <- read_spss("wave2.sav")
```

> Next, recode the variables.

```{r recode}
### Neighborhood, T1
  # recode H1NB6
  wave1$H1NB6[wave1$H1NB6 >5] <- NA
  
  # recode H1NB7
  wave1$NB[wave1$H1NB7 == 1] <- 5
  wave1$NB[wave1$H1NB7 == 2] <- 4
  wave1$NB[wave1$H1NB7 == 4] <- 2
  wave1$NB[wave1$H1NB7 == 5] <- 1
  
  wave1$Neighbor <- apply(wave1[,c("H1NB6", 
                                   "NB")], 1, sum)

### Feeling Scale, T2
  # recode H2FS1
  wave2$H2FS1[wave2$H2FS1 > 3] <- NA
  
  # recode H2FS2
  wave2$H2FS2[wave2$H2FS2 > 3] <- NA
  
  # recode H2FS3
  wave2$H2FS3[wave2$H2FS3 > 3] <- NA
  
  # recode H2FS4
  wave2$FS4[wave2$H2FS4 == 0] <- 3
  wave2$FS4[wave2$H2FS4 == 1] <- 2
  wave2$FS4[wave2$H2FS4 == 2] <- 1
  wave2$FS4[wave2$H2FS4 == 3] <- 0
  
  # recode H2FS5
  wave2$H2FS5[wave2$H2FS5 > 3] <- NA
  
  # recode H2FS6
  wave2$H2FS6[wave2$H2FS6 > 3] <- NA
  
  # recode H2FS7
  wave2$H2FS7[wave2$H2FS7 > 3] <- NA
  
  # recode H2FS8
  wave2$FS8[wave2$H2FS8 == 0] <- 3
  wave2$FS8[wave2$H2FS8 == 1] <- 2
  wave2$FS8[wave2$H2FS8 == 2] <- 1
  wave2$FS8[wave2$H2FS8 == 3] <- 0
  
  # recode H2FS9
  wave2$H2FS9[wave2$H2FS9 > 3] <- NA
  
  # recode H2FS10
  wave2$H2FS10[wave2$H2FS10 > 3] <- NA
  
  # recode H2FS11
  wave2$FS11[wave2$H2FS11 == 0] <- 3
  wave2$FS11[wave2$H2FS11 == 1] <- 2
  wave2$FS11[wave2$H2FS11 == 2] <- 1
  wave2$FS11[wave2$H2FS11 == 3] <- 0
  
  # recode H2FS12
  wave2$H2FS12[wave2$H2FS12 > 3] <- NA
  
  # recode H2FS13
  wave2$H2FS13[wave2$H2FS13 > 3] <- NA
  
  # recode H2FS14
  wave2$H2FS14[wave2$H2FS14 > 3] <- NA
  
  # recode H2FS15
  wave2$FS15[wave2$H2FS15 == 0] <- 3
  wave2$FS15[wave2$H2FS15 == 1] <- 2
  wave2$FS15[wave2$H2FS15 == 2] <- 1
  wave2$FS15[wave2$H2FS15 == 3] <- 0
  
  # recode H2FS16
  wave2$H2FS16[wave2$H2FS16 > 3] <- NA
  
  # recode H2FS17
  wave2$H2FS17[wave2$H2FS17 > 3] <- NA
  
  # recode H2FS18
  wave2$H2FS18[wave2$H2FS18 > 3] <- NA
  
  # recode H2FS19
  wave2$H2FS19[wave2$H2FS19 > 3] <- NA
  
  # calculate depressive symptoms
  wave2$Depression <- apply(wave2[,c("H2FS1", 
                                     "H2FS2", 
                                     "H2FS3", 
                                     "FS4", 
                                     "H2FS5", 
                                     "H2FS6", 
                                     "H2FS7", 
                                     "FS8", 
                                     "H2FS9", 
                                     "H2FS10", 
                                     "FS11", 
                                     "H2FS12", 
                                     "H2FS13", 
                                     "H2FS14", 
                                     "FS15", 
                                     "H2FS16", 
                                     "H2FS17", 
                                     "H2FS18", 
                                     "H2FS19")], 1, sum)
  
### Protective Factors, T1
  # recode H1PR1
  wave1$H1PR1[wave1$H1PR1 == 6] <- NA
  wave1$H1PR1[wave1$H1PR1 > 95] <- NA  
  
  # recode H1PR2
  wave1$H1PR2[wave1$H1PR2 == 6] <- NA
  wave1$H1PR2[wave1$H1PR2 > 95] <- NA  
  
  # recode H1PR3
  wave1$H1PR3[wave1$H1PR3 == 6] <- NA
  wave1$H1PR3[wave1$H1PR3 > 95] <- NA  
  
  # recode H1PR4
  wave1$H1PR4[wave1$H1PR4 == 6] <- NA
  wave1$H1PR4[wave1$H1PR4 > 95] <- NA  
  
  # recode H1PR5
  wave1$H1PR5[wave1$H1PR5 == 6] <- NA
  wave1$H1PR5[wave1$H1PR5 > 95] <- NA  
  
  # recode H1PR8
  wave1$H1PR8[wave1$H1PR8 == 6] <- NA
  wave1$H1PR8[wave1$H1PR8 > 95] <- NA  
  
  # calculate PF
  wave1$PF <- apply(wave1[,c("H1PR1", 
                             "H1PR2", 
                             "H1PR3", 
                             "H1PR4", 
                             "H1PR5", 
                             "H1PR8")], 1, sum)
```

> Now, let's test the hypothesis.

```{r stats}
### make new data sets
  merged <- merge(wave1, wave2, by = 1, all = TRUE)

  # Hypothesis 3
  H3<- subset(merged, complete.cases(merged[, c("Neighbor", 
                                                "Depression", 
                                                "PF")]))
  
    # total effect model
    h3_total_model <- lm(Depression ~ Neighbor, data = H3)
    summary(h3_total_model)
    
    # partial Effect Model
    h3_partial_model <- lm(Depression ~ Neighbor + PF, data = H3)
    summary(h3_partial_model)
```

> The neighborhood environment was found to be a significant longitudinal predictor of mental health. A more positive Wave 1 neighborhood environment predicted lower levels of Wave 2 depressive symptoms (β = -0.56, p < .001), as hypothesized. This effect was not attenuated and remained significant after Wave 1 Protective Factors were included in the model.